Thermodynamic response functions in a cell fluid model

### 2024-09-03 main_morse_response_functions.tex
remove notes
introduce reduced temperature

	The coefficients $g_n$ are given by the formulas:
	\begin{align}
		& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  \nonumber \\
		& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, \quad \qquad  g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, \nonumber\\
		& a_4 = -g_4,
	\end{align}
	where $T_n(p,\alpha^*)$ are the following special functions
	\begin{equation}
		T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n {\rm e}^{-pm^2}.
	\end{equation}
	Here $\alpha^*=v e^{\beta_c\mu^*}$, and the parameter $p$ has the form
	\begin{equation}
		p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
	\end{equation} 
	{\color{Red} {\bf{NOTE 1: }} Quantities $\alpha^*$ has the dimension of volume.}

	The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite[see Eq.(31)]{KozlovskiiDobush2020} it was found that
	\begin{equation*}
		\varepsilon\beta_c = 0.200, \quad \frac{k_{\rm B} T_c}{\varepsilon} = 4.995.
	\end{equation*}
	Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature. The numerical values for other coefficients used in this paper are the same as those in~\cite[see Eqs.(5), (23), and (24)]{KozlovskiiDobush2020}:
	\begin{eqnarray}
		\label{params}
		\chi_0 = 0.07, & \quad \gamma = 1.65, \nonumber\\
		n_0 = 1.521, & \quad R_0/\alpha = 2.9544, \nonumber\\
		\alpha^* = 5.0 & \quad p = 1.0.
	\end{eqnarray}

@@

	Let us introduce the reduced pressure
	\begin{equation}
		\label{def:reduced_pres}
		P^* \equiv \frac{P v}{\varepsilon}.
	\end{equation}
	Considering the equation of state~\eqref{eq:eosMT}, $P^*$ is explicitly written as
	\begin{equation}
		\label{eq:eosPTM_reduced}
		P^* = \frac{1 + \tau}{\varepsilon \beta_c} \left( E_\mu(M, T) + M \bar \rho_0 + \frac{1}{2} \tilde D(0) \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4 \right).
	\end{equation}
	This equation can be easily represented graphically. Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various values of the reduced temperature $\tau$, at and above the critical temperature. At the critical point $M=0, \tau=0$, $P^*_c = 1.606$.

@@

	This dependency can also be expressed explicitly. To achieve this, we combine the equations~\eqref{eq:density} and~\eqref{eq:ro_M} to express the effective chemical potential $M$ as a function of density and temperature:
	\begin{equation}\label{eq:M_nT}
		\bar M = \frac{\rho_{n}}{\beta W(0)} - (\bar n - n_g),
	\end{equation}
	where
	\begin{align} \label{eq:ro_nT}
		 & \rho_{n} = - 2 \left(\frac{g_3^2 - 2g_2 g_4}{g_4^2} \right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \\
		 & \alpha_n = \arccos \left( \left( - \frac{9 g_4^4}{\left( 2 g_2 g_4 - g_3^2\right)^3}\right)^{\frac{1}{2}} (n_g - \bar n)\right). \nonumber 
	\end{align}
	{\color{Red} {\bf{NOTE 2: }} Roma: In my opinion, under {\it{arccos}} we should have 3, not $\frac{3}{2}$ \color{Green} Oksana: Yes, you are right, I've fixed that }

	At $T>T_c$ the equation of state of a cell fluid model in terms of the density and the temperature has the following form
	\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\eta (\bar n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4,
	\end{equation}
 The function $E_\nu (\bar n, T)$ in \eqref{eq:ro_nT} is the function $E_\mu (M,T)$ \eqref{eq:E_mu} rewritten in terms of density and temperature taking into account the expression \eqref{eq:M_nT}. For the reduced pressure we write
 \begin{equation}
 	\label{eq:eosPTn_reduced}
 	P^* = \frac{1 + \tau}{\varepsilon \beta_c} \left(E_\eta (\bar n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4\right).
 \end{equation}

@@

	We introduce the reduced thermal pressure coefficient by
	\begin{equation}
		\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\beta^*_V = \varepsilon\beta_c 
		\left[ \left(\frac{\partial P^*}{\partial \tau}\right)_{\mu} 
		- \left(\frac{\partial P^*}{\partial \mu}\right)_T 
		\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T} 
		\right].
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the thermal pressure coefficient follow immediately from the definition~\eqref{def:therm_pres_coef}
	\begin{equation}
		\beta_V = \left(\frac{\partial P}{\partial T}\right)_\rho
	\end{equation}
	and
	\begin{equation}
		\beta^*_V = \varepsilon\beta_c \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho}
	\end{equation}

@@

	We introduce the reduced thermal expansion coefficient by
	\begin{equation}
		\alpha^*_P = \frac{\varepsilon}{k_{\rm B}} \alpha_P.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\alpha^*_P = \varepsilon \beta_c 
		\left( 
		-\frac{1}{\bar{n}}\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		+ \frac{1}{\bar{n}} \left(\frac{\partial \bar{n}}{\partial \mu}\right)_{T}
		\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
		\right).
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then we need to rewrite the definition for $\alpha_P$ accordingly. We start with~\eqref{eq:alpha_vs_rho} and apply the cyclic relation
	\begin{equation}
		\left(\frac{\partial \rho}{\partial T}\right)_P = - \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T
	\end{equation}
	to get for $\alpha_P$
	\begin{equation}
		\alpha_P = \frac{1}{\rho} \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T.
	\end{equation}
	or for $\alpha^*_P$
	\begin{equation}
		\alpha^*_P = \frac{\varepsilon\beta_c}{\bar{n}} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho} \left(\frac{\partial P^*}{\partial \bar{n}}\right)^{-1}_T.
	\end{equation}

>>>>

	The coefficients $g_n$ are given by the formulas:
	\begin{align}
		& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  \nonumber \\
		& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, \quad \qquad  g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, \nonumber\\
		& a_4 = -g_4,
	\end{align}
	where $T_n(p,\alpha^*)$ are the following special functions
	\begin{equation}
		T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n {\rm e}^{-pm^2}.
	\end{equation}
	Here $\alpha^*=v e^{\beta_c\mu^*}$, and the parameter $p$ has the form
	\begin{equation}
		p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
	\end{equation}

	The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite[see Eq.(31)]{KozlovskiiDobush2020} it was found that
	\begin{equation*}
		\varepsilon\beta_c = 0.200, \quad \frac{k_{\rm B} T_c}{\varepsilon} = 4.995.
	\end{equation*}
	We also use the reduced temperature defined as $T = k_{\rm B}T/\varepsilon$, and its critical value $T^*_c = 4.995.$

	Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature. The numerical values for other coefficients used in this paper are the same as those in~\cite[see Eqs.(5), (23), and (24)]{KozlovskiiDobush2020}:
	\begin{eqnarray}
		\label{params}
		\chi_0 = 0.07, & \quad \gamma = 1.65, \nonumber\\
		n_0 = 1.521, & \quad R_0/\alpha = 2.9544, \nonumber\\
		\alpha^* = 5.0 & \quad p = 1.0.
	\end{eqnarray}

@@

	Let us introduce the reduced pressure
	\begin{equation}
		\label{def:reduced_pres}
		P^* \equiv \frac{P v}{\varepsilon}.
	\end{equation}
	Considering the equation of state~\eqref{eq:eosMT}, $P^*$ is explicitly written as
	\begin{equation}
		\label{eq:eosPTM_reduced}
		P^* = (1 + \tau)T^*_c \left( E_\mu(M, T) + M \bar \rho_0 + \frac{1}{2} \tilde D(0) \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4 \right).
	\end{equation}
	This equation can be easily represented graphically. Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various values of the reduced temperature $\tau$, at and above the critical temperature. At the critical point $M=0, \tau=0$, $P^*_c = 1.606$.

@@

	This dependency can also be expressed explicitly. To achieve this, we combine the equations~\eqref{eq:density} and~\eqref{eq:ro_M} to express the effective chemical potential $M$ as a function of density and temperature:
	\begin{equation}\label{eq:M_nT}
		\bar M = \frac{\rho_{n}}{\beta W(0)} - (\bar n - n_g),
	\end{equation}
	where
	\begin{align} \label{eq:ro_nT}
		 & \rho_{n} = - 2 \left(\frac{g_3^2 - 2g_2 g_4}{g_4^2} \right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \\
		 & \alpha_n = \arccos \left( \left( - \frac{9 g_4^4}{\left( 2 g_2 g_4 - g_3^2\right)^3}\right)^{\frac{1}{2}} (n_g - \bar n)\right). \nonumber 
	\end{align}

	At $T>T_c$ the equation of state of a cell fluid model in terms of the density and the temperature has the following form
	\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\eta (\bar n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4,
	\end{equation}
 The function $E_\nu (\bar n, T)$ in \eqref{eq:ro_nT} is the function $E_\mu (M,T)$ \eqref{eq:E_mu} rewritten in terms of density and temperature taking into account the expression \eqref{eq:M_nT}. For the reduced pressure we write
 \begin{equation}
 	\label{eq:eosPTn_reduced}
 	P^* = (1 + \tau)T^*_c \left(E_\eta (\bar{n},T) + \bar{M} \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4\right).
 \end{equation}

@@

	We introduce the reduced thermal pressure coefficient by
	\begin{equation}
		\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\beta^*_V = \frac{1}{T^*_c} 
		\left[ \left(\frac{\partial P^*}{\partial \tau}\right)_{\mu} 
		- \left(\frac{\partial P^*}{\partial \mu}\right)_T 
		\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T} 
		\right].
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the thermal pressure coefficient follow immediately from the definition~\eqref{def:therm_pres_coef}
	\begin{equation}
		\beta_V = \left(\frac{\partial P}{\partial T}\right)_\rho
	\end{equation}
	and
	\begin{equation}
		\beta^*_V = \frac{1}{T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho}
	\end{equation}

@@

	We introduce the reduced thermal expansion coefficient by
	\begin{equation}
		\alpha^*_P = \frac{\varepsilon}{k_{\rm B}} \alpha_P.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\alpha^*_P = \frac{1}{T^*_c}
		\left( 
		-\frac{1}{\bar{n}}\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		+ \frac{1}{\bar{n}} \left(\frac{\partial \bar{n}}{\partial \mu}\right)_{T}
		\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
		\right).
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then we need to rewrite the definition for $\alpha_P$ accordingly. We start with~\eqref{eq:alpha_vs_rho} and apply the cyclic relation
	\begin{equation}
		\left(\frac{\partial \rho}{\partial T}\right)_P = - \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T
	\end{equation}
	to get for $\alpha_P$
	\begin{equation}
		\alpha_P = \frac{1}{\rho} \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T.
	\end{equation}
	or for $\alpha^*_P$
	\begin{equation}
		\alpha^*_P = \frac{1}{\bar{n} T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho} \left(\frac{\partial P^*}{\partial \bar{n}}\right)^{-1}_T.
	\end{equation}
